---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Load the activity data and take a look at first 6 records and check data attributes.

```{r}
activity <- read.csv('activity.csv')
head(activity)
str(activity)
```

## What is mean total number of steps taken per day?

```{r}
totalsteps <- tapply(activity$steps,as.Date(activity$date),sum,na.rm=T)
hist(totalsteps,breaks=50,main='Total numbers of steps taken per day',xlab='Total Steps',ylab='Number of Days',col='cadetblue3')
```

The mean total number of steps per day is: `r prettyNum(mean(totalsteps),big.mark = ',')`.

The median total number of steps per day is: `r prettyNum(median(totalsteps),big.mark = ',')`.

## What is the average daily activity pattern?
```{r}
avgsteps <- aggregate(activity$steps~activity$interval,FUN = mean,na.rm=T)
colnames(avgsteps) <- c('interval','steps')
with(avgsteps,plot(interval,steps,type='l',ylab='Daily Average Steps',xlab='5 Minutes Interval',main='Average Daily Activity Pattern'))
max_int <- avgsteps[which.max(avgsteps$steps),'interval']
```

The 5-minutes interval which contains the maximum number of steps, on average across all the days in the datasetis: `r max_int`.

## Imputing missing values

There are `r sum(is.na(activity$steps))` missing values in the dataset. We are imputing missing values with the mean. Below histogram shows the total number of steps taken each day after we imputing missing values.

```{r message=FALSE}
library(Hmisc)
library(ggplot2)
```
```{r}
activity2 <- activity
activity2$steps <- impute(activity$steps, fun=mean)
totalsteps2 <- tapply(activity2$steps,as.Date(activity2$date),sum)
hist(totalsteps2,breaks=50,main='Total numbers of steps taken per day',xlab='Total Steps',ylab='Number of Days',col='cadetblue3')
```

After imputing missing values with mean, the mean total number of steps per day is: `r prettyNum(mean(totalsteps2),big.mark = ',')`.

After imputing missing values with mean, the median total number of steps per day is: `r prettyNum(median(totalsteps2),big.mark = ',')`.

The values are higher than what we reported before. 

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity2$date <- as.Date(activity2$date)
activity2$wday<- ifelse(weekdays(activity2$date) %in% c('Saturday','Sunday'),'Weekend','Weekday')
avgsteps2 <- aggregate(activity2$steps~activity2$interval + activity2$wday,FUN = mean)
colnames(avgsteps2) <- c('interval','wday','steps')

g <- ggplot(avgsteps2,aes(interval,steps))
g + geom_line() + labs(x="5-minute Interval",y="Avarage Number of Steps") + facet_grid(wday~.)
```

Looking at the plot for activity pattern between weekdays and weekneds, we can see that the activity is lower during weekends.
